---
title: "Salaries_Presentation_Slides"
name:
format: revealjs
---

```{r}
#/ echo: false
#/ warning: false
#/ output:false
library(tidyverse)

theme_set(theme_minimal())

salary <- read.csv("data/v7_Latest_Data_Science_Salaries.csv", stringsAsFactors = TRUE) |>
  filter(!is.na(Salary.in.USD))

## Question: What happens in the case_when when there is a "Business Intelligence Analyst"?

salary = salary |>
  mutate(Job.Category = case_when(str_detect(Job.Title, pattern = "AI ") == TRUE ~ "Artificial Intelligence",
                                  str_detect(Job.Title, pattern = "Autonomous") == TRUE ~ "Artificial Intelligence",
                                  str_detect(Job.Title, pattern = "Computer Vision") == TRUE ~ "Artificial Intelligence",
                                  str_detect(Job.Title, pattern = "NLP") == TRUE ~ "Artificial Intelligence",
                                  str_detect(Job.Title, pattern = "Prompt Engineer") == TRUE ~ "Artificial Intelligence",
                                  str_detect(Job.Title, pattern = "BI ") == TRUE ~ "Business Intelligence",
                                  str_detect(Job.Title, pattern = "Business") == TRUE ~ "Business Intelligence",
                                  str_detect(Job.Title, pattern = "Data Product") == TRUE ~ "Business Intelligence",
                                  str_detect(Job.Title, pattern = "Financ") == TRUE ~ "Business Intelligence",
                                  str_detect(Job.Title, pattern = "Data Sci") == TRUE ~ "Data Science",
                                  str_detect(Job.Title, pattern = "Data Lead") == TRUE ~ "Data Science",
                                  str_detect(Job.Title, pattern = "Data Manage") == TRUE ~ "Data Science",
                                  str_detect(Job.Title, pattern = "Data Specialist") == TRUE ~ "Data Science",
                                  str_detect(Job.Title, pattern = "Head of Data") == TRUE ~ "Data Science",
                                  str_detect(Job.Title, pattern = "ML") == TRUE ~ "Machine Learning",
                                  str_detect(Job.Title, pattern = "Machine") == TRUE ~ "Machine Learning",
                                  str_detect(Job.Title, pattern = "Deep Learning") == TRUE ~ "Machine Learning",
                                  str_detect(Job.Title, pattern = "Big ") == TRUE ~ "Big Data",
                                  str_detect(Job.Title, pattern = "Cloud") == TRUE ~ "Cloud Data",
                                  str_detect(Job.Title, pattern = "Data Architect") == TRUE ~ "Data Organization",
                                  str_detect(Job.Title, pattern = "Data Engineer") == TRUE ~ "Data Organization",
                                  str_detect(Job.Title, pattern = "Data Developer") == TRUE ~ "Data Organization",
                                  str_detect(Job.Title, pattern = "Data Infrastructure") == TRUE ~ "Data Organization",
                                  str_detect(Job.Title, pattern = "ETL ") == TRUE ~ "Data Organization",
                                  str_detect(Job.Title, pattern = "Data Operations") == TRUE ~ "Data Operations",
                                  str_detect(Job.Title, pattern = "DevOps") == TRUE ~ "Data Operations",
                                  str_detect(Job.Title, pattern = "Data Integration") == TRUE ~ "Data Operations",
                                  str_detect(Job.Title, pattern = "Data Strateg") == TRUE ~ "Data Operations",
                                  str_detect(Job.Title, pattern = "Decision Scientist") == TRUE ~ "Data Operations",
                                  str_detect(Job.Title, pattern = "Data Model") == TRUE ~ "Data Modeling",
                                  str_detect(Job.Title, pattern = "Data Quality") == TRUE ~ "Data Quality",
                                  str_detect(Job.Title, pattern = "Data Vis") == TRUE ~ "Data Visualization",
                                  str_detect(Job.Title, pattern = "Research Analyst") == TRUE ~ "Research",
                                  str_detect(Job.Title, pattern = "Research Engineer") == TRUE ~ "Research",
                                  str_detect(Job.Title, pattern = "Research Scientist") == TRUE ~ "Research",
                                  str_detect(Job.Title, pattern = "Applied Scientist") == TRUE ~ "Research",
                                  str_detect(Job.Title, pattern = "Analy") == TRUE ~ "Data Analytics"
                                  )) |>
  mutate(Company.Location = as.character(Company.Location),
         Employment.Type = as.character(Employment.Type),
         Experience.Level = fct_relevel(Experience.Level, c("Entry", "Mid", "Senior", "Executive")),
         Company.Size = fct_relevel(Company.Size, c("Small", "Medium", "Large")))
```


## Data

* I found this dataset, called Latest Data Science Salaries, published by Sourav Banerjee
* The data ranges from 2020 to the beginning of 2024 and is primarily of jobs in the USA, but also has some global data as well
  
## Variables of Interest

* `Salary.in.USD`: job salary converted to USD
* `Job.Title`: job position title
* `Job.Category`: category of job position (added by myself)
* `Experience.Level`: factor of individual's experience level (Entry, Mid, Senior, Executive)
* `Company.Size`: factor of the company's size (Small, Medium, Large)
* `Company.Location`: geographic location of the company
* `Employment.Type`: factor of job position type (Full-Time, Part-Time, Freelance)
* `Year`: year of employment (2020-2024)

## Questions of Interest:

1. What is the distribution of salary for Data Scientist Entry- and Mid-level positions?
2. How has the experience level and pay of Data Analysts changed over the course of this dataset (2020 through early 2024)?
3. For the Engineer position, how does the shape of the pay differ for different specialties (AI Engineer vs. Data Science Engineer vs. Machine Learning Engineer)?

## Visual 1

```{r}
ds_static = salary |> 
  filter(Job.Title == "Data Scientist") |>
  filter(Experience.Level %in% c("Entry", "Mid"))
ggplot(data = ds_static, aes(x = Salary.in.USD)) +
      geom_histogram(bins = 15, col = "lightblue4", fill = "lightblue") +
      labs(title = glue::glue("Histogram of Salary (USD) for Entry- and Mid-Level Data Scientist Jobs")) +
      theme_minimal(base_size = 20) +
      labs(x = "Salary (USD)")
```

## Visual 2

```{r}
da_static = salary |>
      filter(Job.Title == "Data Analyst")
ggplot(data = da_static, aes(x = Salary.in.USD, color = Experience.Level)) +
      geom_freqpoly(linewidth = 1) +
      facet_wrap(~ Year) +
      labs(title = glue::glue("Frequency Plot of Salary (USD) for Data Analyst Jobs")) +
      theme_minimal(base_size = 20) +
      labs(x = "Salary (USD)", color = "Experience Level") +
      scale_color_viridis_d() +
      theme(axis.text.x = element_text(angle = -45, hjust = 1))
```

## Visual 3

```{r}
violin_df = salary |>
      filter(Job.Title %in% c("AI Engineer", "Data Science Engineer", "Machine Learning Engineer"))
ggplot(data = violin_df, aes(x = Job.Title, y = Salary.in.USD)) +
      geom_violin(fill = "lightgreen", color = "forestgreen") +
      labs(title = glue::glue("Violin Plot of Salary (USD)"), x = "Job Title", y = "Salary (USD)") +
      theme_minimal(base_size = 20)
```

